Automatic prostate cancer detection through \acs*{dce}-\acs*{mri} images: all you need is a good normalization

### 2017-01-27 03_experiments/experiments.tex
Advance the correction

Parameter estimation of the quantification methods are related to fit a specific model to the \ac{dce}-\ac{mri} data.
Therefore, this section report the goodness of fitting by computing the coefficient of determination $R^2$ such as in Eq.\,\eqref{eq:r2}

@@

Mean and standard-deviation of the coefficient of determination $R^{2}$ is reported in Table~\ref{tab:r2} for each quantification model.
Brix, Hoffmann, and Tofts models are fitted with a coefficient $R^{2}$ superior to 0.80.
Additionally, the proposed \ac{pun} model does not seem to fit well the data.
Data normalization improves the coefficient $R^2$ for all the methods apart of the Hoffmann model.
The large standard deviation for this model might imply that there are some cases where the fitting fails.

\subsection{Detection of \acs*{cap} using pharmacokinetic parameters}

To study the potential benefit of our normalization, \ac{cap} are detected at a voxel level using pharmacokinetic parameters estimated from un-normalized and normalized \ac{dce}-\ac{mri} data.
Each individual pharmacokinetic parameter is classified to evaluate their individual discriminative power to detect \ac{cap}.
Therefore, a \ac{rf} classifier is used in conjunction with a \ac{lopo}.
The use of \ac{rf} is motivated since that it leads to the best performance in the state-of-the-art methods~\citep{litjens2014computer,lemaitre2015computer}.
Results are summarized in Table~\ref{tab:resfeats} in terms of \ac{auc}.
Normalization can improve the detection of \ac{cap}; however, the benefit of normalization is more obvious by combining together the pharmacokinetic features of a given model (e.g., $A$, $k_ep$, and $k_el$ for Brix model), as previously done in traditional \ac{cad} system~\citep{lemaitre2015computer}.
For the latter configuration, results are summarized by performing a \ac{roc} analysis and computing the \ac{auc}, as reported in Fig.\,\ref{fig:normpharmarf}.
Quantification using normalized data outperforms quantification using un-normalized data in terms of classification performance apart of Hoffmann and Tofts population-based \ac{aif} models.
The reasons behind the decrease of the \ac{auc} might be related to: (i) a poor fitting as discussed in Sect.\,\ref{sec:fit} (cf., Hoffmann model) and (ii) a small number of patients while estimating some parameters (cf., Tofts model).
The best classification performance are obtained using the semi-quantitative approach with an \ac{auc} of 0.655.

\subsection{Classification of the entire enhanced \acs*{dce}-\acs*{mri} signal}

As stated in the introduction, the quantification methods are extracting a set of parameters characterizing the enhancement \ac{dce}-\ac{mri} signal.
However, this extraction might lead to a loss of information.
This experiment is performed to assess if making use of the whole \ac{dce}-\ac{mri} signal instead of the just the pharmacokinetic parameters can improve the classification performance.
Therefore, each enhanced \ac{dce}-\ac{mri} signal, normalized and un-normalized, is classified using a \ac{rf} classifier in a \ac{lopo} fashion.
The \ac{roc} analysis and \ac{auc} are reported in Fig.\,\ref{fig:rfnormdcesignal}.
Classification without normalization lead to the worst performance, with an \ac{auc} of 0.568.
However, data normalization in conjunction with the use of the whole \ac{dce}-\ac{mri} signal is the strategy which outperforms all others, with an \ac{auc} of 0.666.

>>>>

\begin{table*}
  \caption{Coefficient of determination $R^{2}$ (i.e., $\mu \ (\pm \sigma)$), while fitting data with the different quantification models.}
  \centering
  \resizebox{\columnwidth}{!}{
  \begin{tabular}{lc c c c c c}
    \toprule
    Data type & Brix & Hoffmann & Tofts population \ac{aif} & Tofts patient \ac{aif} & \ac{pun} & Semi-quantitative \\
    \midrule
    Un-normalized & $0.85 \ (\pm 0.11)$ & $0.81 \ (\pm 0.17)$ & $0.84 \ (\pm 0.14)$ & $0.88 \ (\pm 0.12)$ & $0.27 \ (\pm 0.18)$ & $0.64 \ (\pm 0.24)$  \\
    Normalized    & $0.92 \ (\pm 0.05)$ & $0.72 \ (\pm 0.32)$ & $0.92 \ (\pm 0.06)$ & $0.90 \ (\pm 0.10)$ & $0.28 \ (\pm 0.20)$ & $0.75 \ (\pm 0.20)$  \\
    \bottomrule
  \end{tabular}
  }
  \label{tab:r2}
\end{table*}

Parameter estimation of the quantification methods are related to fit a specific model to the \ac{dce}-\ac{mri} data.
Therefore, this section reports the goodness of fitting by computing the coefficient of determination $R^2$ such as in Eq.\,\eqref{eq:r2}

@@

Mean and standard-deviation of the coefficient of determination $R^{2}$ is reported in Table~\ref{tab:r2} for each quantification model.
Brix, Hoffmann, and Tofts models are fitted with a coefficient $R^{2}$ superior to 0.80.
Additionally, the proposed \ac{pun} model does not to fit well the data.
After introspection of the fitted curves, the original model --- as
formulated in Eq.\,\eqref{eq:pun} --- does not provide enough degrees
of freedom to fit well the data. Additional parameters should be
integrated in this model to control the translation and amplitude of
the model to be fitted.
Data normalization improves the coefficient $R^2$ for all the methods apart of the Hoffmann model.
The large standard deviation for this model might imply that there are some cases where the fitting fails.

% \subsection{Detection of \acs*{cap} using pharmacokinetic parameters}

\subsection{\acs*{cap} detection using pharcokinetic,
  semi-quantitative, and entire enhanced signal}

\begin{table*}
  \caption{\acs*{auc} (i.e., $\mu \ (\pm \sigma)$) for each individual pharmacokinetic parameter using a \acs*{rf} classifier.}
  \centering
  \resizebox{\columnwidth}{!}{
  \begin{tabular}{l c c}
    \toprule
    \textbf{Features} & Un-normalized data & Normalized data \\
    \midrule
    \textbf{Brix model} & & \\
    \quad $A$         & $0.540\ (\pm 0.069)$ & $0.555\ (\pm 0.080)$ \\
    \quad $k_{el}$    & $0.549\ (\pm 0.062)$ & $0.577\ (\pm 0.093)$ \\
    \quad $k_{ep}$    & $0.506\ (\pm 0.032)$ & $0.497\ (\pm 0.019)$ \\
    \textbf{Hoffmann model} & & \\
    \quad $A$         & $0.516\ (\pm 0.020)$ & $0.508\ (\pm 0.031)$ \\
    \quad $k_{el}$    & $0.545\ (\pm 0.066)$ & $0.529\ (\pm 0.065)$ \\
    \quad $k_{ep}$    & $0.550\ (\pm 0.063)$ & $0.545\ (\pm 0.060)$ \\
    \textbf{Tofts model with population \ac{aif}} & & \\
    \quad $K_{trans}$ & $0.556\ (\pm 0.086)$ & $0.565\ (\pm 0.097)$ \\
    \quad $k_{ep}$    & $0.506\ (\pm 0.026)$ & $0.528\ (\pm 0.038)$ \\
    \quad $v_{p}$     & $0.533\ (\pm 0.064)$ & $0.548\ (\pm 0.082)$ \\
    \textbf{Tofts model with patient \ac{aif}} & & \\
    \quad $K_{trans}$ & $0.563\ (\pm 0.077)$ & $0.548\ (\pm 0.060)$ \\
    \quad $k_{ep}$    & $0.492\ (\pm 0.025)$ & $0.491\ (\pm 0.020)$ \\
    \quad $v_{p}$     & $0.530\ (\pm 0.069)$ & $0.495\ (\pm 0.033)$ \\
    \textbf{\ac{pun} model} & & \\
    \quad $a_0$       & $0.521\ (\pm 0.040)$ & $0.530\ (\pm 0.045)$ \\
    \quad $r$         & $0.550\ (\pm 0.085)$ & $0.573\ (\pm 0.097)$ \\
    \quad $\beta$     & $0.531\ (\pm 0.051)$ & $0.549\ (\pm 0.068)$ \\
    \textbf{Semi-quantitative analysis} & & \\
    \quad wash-in     & $0.587\ (\pm 0.107)$ & $0.533\ (\pm 0.032)$ \\
    \quad wash-out    & $0.516\ (\pm 0.037)$ & $0.486\ (\pm 0.035)$ \\
    \quad IAUC        & $0.506\ (\pm 0.048)$ & $0.513\ (\pm 0.032)$ \\
    \quad $\tau$      & $0.565\ (\pm 0.104)$ & $0.537\ (\pm 0.089)$ \\
    \quad $S_M - S_0$ & $0.560\ (\pm 0.083)$ & $0.532\ (\pm 0.029)$ \\
    \bottomrule
  \end{tabular}
  }
  \label{tab:resfeats}
\end{table*}

\begin{figure*}
  \centering
  \hspace*{\fill}
  \subfigure[Quantiative model without normalization.]{\label{fig:rfpharmaunorm}\includegraphics[width=.49\textwidth]{03_experiments/figures/unormalized/unormalized_methods_0.pdf}} \hfill
  \subfigure[Quantitative model With normalization.]{\label{fig:rfpharmanorm}\includegraphics[width=.49\textwidth]{03_experiments/figures/normalized/normalized_methods_0.pdf}}
  \hspace*{\fill}
  \hspace*{\fill}
  \subfigure[Entire enhanced \acs*{dce} signal classification.]{\label{fig:rfnormdcesignal}\includegraphics[width=.49\textwidth]{03_experiments/figures/full_signal_0.pdf}}
  \hspace*{\fill}
  \caption{\acs*{roc} analysis using a \acs*{rf} classifier with and
    without normalization using different approaches: \protect\subref{fig:rfpharmaunorm} -
    \protect\subref{fig:rfpharmanorm} pharmacokinetic and
    semi-quantitative models without and with normalization,
    respectively; \protect\subref{fig:rfnormdcesignal} entire enhanced
  \acs*{dce}-\acs*{mri} signal.}
  \label{fig:normpharmarf}
\end{figure*}

\begin{figure}
  \centering
  \includegraphics[width=1.\linewidth]{03_experiments/figures/wilcoxon.pdf}
  \caption{Wilcoxon signed-ranked test to compare each pair of
    classifiers. The annotation in the matrix corresponds to the
    p-values ($p$). $p < 0.05$ indicates that a pair of classifiers
    are significantly different, consequently the classifier with the
    highest \acs*{auc} significantly outperforms the other
    one. In the co-occurrence matrix, {\color{blue}blue} cells
    correspond to lower $p$ while {\color{red} red} cells are
    representing greater $p$.}
  \label{fig:wilcoxon}
\end{figure}

To study the potential benefit of our normalization, \ac{cap} are detected at a voxel level using pharmacokinetic parameters estimated from un-normalized and normalized \ac{dce}-\ac{mri} data.
Each individual pharmacokinetic parameter is classified to evaluate their individual discriminative power to detect \ac{cap}.
Therefore, a \ac{rf} classifier is used in conjunction with a \ac{lopo}.
The use of \ac{rf} is motivated since that it leads to the best performance in the state-of-the-art methods~\citep{litjens2014computer,lemaitre2015computer}.
Results are summarized in Table~\ref{tab:resfeats} in terms of \ac{auc}.
Normalization can improve the detection of \ac{cap}; however, the benefit of normalization is more obvious by combining together the pharmacokinetic features of a given model (e.g., $A$, $k_ep$, and $k_el$ for Brix model), as previously done in traditional \ac{cad} system~\citep{lemaitre2015computer}.
For the latter configuration, results are summarized by performing a \ac{roc} analysis and computing the \ac{auc}, as reported in Fig.\,\ref{fig:normpharmarf}.
Quantification using normalized data outperforms quantification using un-normalized data in terms of classification performance apart of Hoffmann and Tofts population-based \ac{aif} models.
The reasons behind the decrease of the \ac{auc} might be related to: (i) a poor fitting as discussed in Sect.\,\ref{sec:fit} (cf., Hoffmann model) and (ii) a small number of patients while estimating some parameters (cf., Tofts model).
The best classification performance are obtained using the semi-quantitative approach with an \ac{auc} of $0.655$.

%\subsection{Classification of the entire enhanced \acs*{dce}-\acs*{mri} signal}

\begin{figure*}
  \centering
  \includegraphics[width=1.\linewidth]{03_experiments/figures/plot_auc_by_zones.pdf}
  \caption{Comparison of the distribution of the \acs*{auc} score for
    each prostate zone, with and without normalization.}
  \label{fig:boxplot}
\end{figure*}

As stated in the introduction, the quantification methods are extracting a set of parameters characterizing the enhancement \ac{dce}-\ac{mri} signal.
However, this extraction might lead to a loss of information.
This experiment is performed to assess if making use of the whole \ac{dce}-\ac{mri} signal instead of the just the pharmacokinetic parameters can improve the classification performance.
Therefore, each enhanced \ac{dce}-\ac{mri} signal, normalized and un-normalized, is classified using a \ac{rf} classifier in a \ac{lopo} fashion.
The \ac{roc} analysis and \ac{auc} are reported in Fig.\,\ref{fig:rfnormdcesignal}.
Classification without normalization lead to the worst performance, with an \ac{auc} of $0.568$.
However, data normalization in conjunction with the use of the entire \ac{dce}-\ac{mri} signal is the strategy which outperforms others, with an \ac{auc} of $0.666$.
Comparable outcomes are visible while analyzing the \ac{auc}
distributions scores for each prostate zone. The \ac{auc} score
related to \ac{cap} in \ac{cg} improves using normalized data, for
most of the models.

### 2017-01-27 04_discussions/discussions.tex
Advance the correction

The best classification performance is reached by normalizing the \ac{dce}-\ac{mri} data and use the whole enhanced signal as feature, emphasizing the fact that a loss of information while extracting quantitative parameters.
Furthermore, this normalization is a less complex process than all quantification methods.
However, this strategy suffers from one drawback: the training time of the \ac{rf} classifier increases since that from 3 to 5 features, the feature space becomes a 40 dimensions space.

Nevertheless, this study is performed on a small cohort of patients using a single \ac{mri} machine.
Generalizing the results of this study on a larger dataset acquired from different commercial systems have to be considered to study the robustness of the proposed approach.

>>>>

The best classification performance is reached by normalizing the \ac{dce}-\ac{mri} data and use the entire enhanced signal as feature, emphasizing the fact that a \emph{loss of information} while extracting quantitative parameters.
Furthermore, this normalization is a less complex process than all
quantification methods and significantly improves the classification
performance of the semi-quantitative approach ($p=0.001$) proposed by
\citeauthor{huisman2001accurate}.
Additionally, our normalization improves the \ac{auc} score related to
the detection of \ac{cap} in \ac{cg}, also known to be the most
challenging cases in the diagnosis of \ac{cap}.

However, using the entire enhanced signal in conjunction with the
normalization suffers from one drawback: the training time of the
\ac{rf} classifier increases; instead of using $3$ to $5$ features, the
feature space becomes a $40$ dimensions space. The extraction of the
semi-quantitative parameters leads to a comparable classification performance ---
$0.655$ vs. $0.666$ --- and should be chosen as the alternative method to
consider if the number of feature in the classification is
critical. Despite this fact, the benefit of the normalization method
has been shown for both methods.

Nevertheless, this study is performed on a small cohort of patients using a single \ac{mri} machine.
Generalizing the results of this study on a larger dataset acquired from different commercial systems have to be considered to study the robustness of the proposed approach.
Additionally, our method being non-parametric should provide an
adequate framework to process \ac{dce} data from different
institutions, scanners with different settings.
